---
title: 'Lab 3: Binary Logisitc Regression'
author: "Sachiko Lamen"
date: "2/2/2022"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning= FALSE, message = FALSE)

library(tidyverse)
library(palmerpenguins)
library(GGally)
library(broom)
library(jtools)
library(caret) # use to automate cross validation
library(AICcmodavg)
```

Binary logistic regression - two mutually exclusive outcomes
Multinomial logistic - more than two mutually exclusive outcomes
ordinal logistic regression - ordinal categorical variable (ex low medium high)

In this lab we will find the probabilities that a given penguin falls into Adelie or Chinstrap based on predictor variables (characteristics)

## Binary Logistic Regression
What is the probability it is Adelie? What is the probability it is Chinstrap?

##### Explore data using `GGally::ggpairs`

```{r}
penguins %>%
  select(species, bill_length_mm:sex) %>%
  ggpairs(aes(color = species))
```

##### Make a subset of Chinstrap and Adelie
> class(penguins$species)
[1] "factor"
> levels(penguins$species)
[1] "Adelie"    "Chinstrap" "Gentoo"  

```{r}
adelie_chinstrap <- penguins %>%
  filter(species %in% c("Adelie", "Chinstrap")) %>%
  mutate(species = fct_drop(species)) %>% # drops gentoo completely
  select(-island, -year) %>%
  drop_na()
```
> levels(adelie_chinstrap$species)
[1] "Adelie"    "Chinstrap"

Let's check some trends for variables for these two species
```{r}
ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = sex)) +
  facet_wrap(~species)

ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(aes(color = sex)) +
  facet_wrap(~species)
```

#### Logisitc Regression in R

```{r}
# first write a formula for the regression (we will use again later for cross validation)
f1 <- species ~ body_mass_g + flipper_length_mm + sex

# use glm() generalized linear model -- set to binomial
ad_chin_blr1 <- glm(formula = f1,
                    data = adelie_chinstrap,
                    family = "binomial")
```

look at the results:
```{r}
ad_chin_blr1
summary(ad_chin_blr1)
# for every gram of additional body mass it makes it slightly less likely that the penguin is a Chinstrip
# for every addition mm of flipper length it makes it slightly more likely that the penguin is a chinstrap
# if your species is male compared to the reference level of female, there is a slight decrease that the penguin is a chinstrap ** this doesnt exactly make sense
```
> levels(adelie_chinstrap$species)
[1] "Adelie"    "Chinstrap"
^^ this is important because it tells you the reference level. Adelie is the reference level because it comes first! Adelie coded as 0, Chinstrap coded as 1. 

```{r}
# create tidy data frame of summary statistics, use to compare to initial `GGally::ggpairs` graph
blr1_tidy <- broom::tidy(ad_chin_blr1)
```

```{r}
ggplot(data = adelie_chinstrap, aes(x = species, y = body_mass_g)) +
  geom_jitter(aes(color = sex))

ggplot(data = adelie_chinstrap, aes(x = species, y = flipper_length_mm)) +
  geom_jitter(aes(color = sex))
```
Changes in log odds are difficult to interpret, so we want to convert it into a actual probability...
```{r}
blr1_fitted <- ad_chin_blr1 %>%
  broom::augment(type.predict = "response") #instead of predicting log odds, take them and convert into probability

# look at .fitted column values are probability
```


Lets plot the probabilities based on flipper length
```{r}
ggplot(data = blr1_fitted, aes(x = flipper_length_mm, y = .fitted)) +
  geom_point(aes(color = sex, shape = species)) + # use shape to see if graph is accurately predicting species probability
  geom_smooth(aes(color = sex), se = FALSE) + # add general trendline, remove std error bars
  labs(x = "Flipper Length (mm)",
       y = "Probability of outcome 'Chinstrap'")

# as flipper length gets smaller predicts low probability that pengins are chinstrap, at long flipper lengths, predicts higher probability of chinstrap
```

Visualize model outcomes using `jtools::effect_plot()` shows how prediciton varies based on given variable

```{r}
effect_plot(ad_chin_blr1,
            pred = flipper_length_mm,
            interval = TRUE,
            y.label = "Probability of Chinstrap")

# effect of flipper length on the probability of chinstrap. Lower values of flipper length gnerally will lead to predicitons of adelie, higher will lead to predictions of chinstrap

effect_plot(ad_chin_blr1,
            pred = body_mass_g,
            interval = TRUE,
            y.label = "Probability of Chinstrap")
```

##### Predict species based on new values

What is the probability that a female penguin with 3410 g of body mass and flipper length of 192 mm will be "Chinstrap"?

```{r}
ex1 <- predict(ad_chin_blr1,
               data.frame(sex = "female",
                          body_mass_g = 3410,
                          flipper_length_mm = 192),
               type = "response") # convert log odds into probability

# result = 0.403 --- 40% chance this penguin is a chinstrap
```

Lets do a prediction for multiple penguins!
```{r}
new_df <- data.frame(
  sex = c("male", "male", "female"),
  body_mass_g = c(3298, 4100, 3600),
  flipper_length_mm = c(212, 175, 180)
)

ex2 <- predict(ad_chin_blr1,
               new_df,
               type = "response")
```


#### New model based on bill length

```{r}
f2 <- species ~ bill_length_mm + body_mass_g

ad_chin_blr2 <- glm(formula = f2,
                    data = adelie_chinstrap,
                    family = "binomial")
```

```{r}
# look at the model coefficients and AIC
ad_chin_blr2
summary(ad_chin_blr2)
```
```{r}
ggplot(data = adelie_chinstrap, aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species))
```
```{r}
effect_plot(ad_chin_blr2,
            pred = bill_length_mm,
            interval = TRUE,
            y.label = "Probability of 'Chinstrap'")

effect_plot(ad_chin_blr2,
            pred = body_mass_g,
            interval = TRUE,
            y.label = "Probability of 'Chinstrap'")

# these two graphs make it look like this model has a pretty sure prediction of adelie v chinstrap
```

### Compare the two models : Model Selection

Look at AIC Values
```{r}
AICcmodavg::aictab(list(ad_chin_blr1, ad_chin_blr2))

# says model 2 is better and model 1 is much worse. ∆AIC is so big meaning model 1 is way worse.
```

#### Lets do 10-fold cross validation 
Use predition accuracy as our metric

```{r}
set.seed(123)

n_folds <- 10
folds <- rep(1:n_folds, length.out = nrow(adelie_chinstrap)) # length.out creates a vector of same length as dataframe

ad_chin_kfold <- adelie_chinstrap %>%
  mutate(fold = sample(folds, size = n(), replace = FALSE))
```

```{r}
# create function where x = known value and y = predicted value
pred_acc <- function(x,y){
  accurate <- ifelse(x == y, 1,0) # take string of x's and compare to string of y's and say "this one matches if theres a 1, or it doesnt match if there's a 0
  return(mean(accurate, na.rm = TRUE))
}
```

```{r}
results_df <- data.frame()

for(i in 1:n_folds) {
  kfold_test <- ad_chin_kfold %>%
    filter(fold == i) # fold is equal to the current iteration "i"
  kfold_train <- ad_chin_kfold %>%
    filter(fold != i)
  
  kfold_blr1 <- glm(f1, data = kfold_train, family = "binomial")
  kfold_blr2 <- glm(f2, data = kfold_train, family = "binomial")
  
  kfold_pred <- kfold_test %>%
    mutate(blr1 = predict(kfold_blr1, kfold_test, type = "response"),
           blr2 = predict(kfold_blr2, kfold_test, type = "response")) %>%
    mutate(pred1 = ifelse(blr1 > .50, "Chinstrap", "Adelie"),
           pred2 = ifelse(blr2 > .50, "Chinstrap", "Adelie"))
  kfold_accuracy <- kfold_pred %>%
    summarize(blr1_ac = pred_acc(species, pred1),
              blr2_ac = pred_acc(species, pred2))
  
  results_df <- bind_rows(results_df, kfold_accuracy)
}

results_df %>%
  summarize(blr1_acc = mean(blr1_ac),
            blr2_acc = mean(blr2_ac))
```

Use the `caret` package (classification and regression training) to automate the cross-validation training
```{r}
set.seed(123)

tr_ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 10)

## train the model to do cross validation on 10 folds on adelie_chinstrap dataset 
model1 <- train(f1, data = adelie_chinstrap,
                method = 'glm', family = "binomial", 
                trControl = tr_ctrl)
model1

model2 <- train(f2, data = adelie_chinstrap,
                method = 'glm', family = "binomial", 
                trControl = tr_ctrl)
model2
```


























