---
title: 'ESM 244: Lab 5 - PCA'
author: "Sachiko Lamen"
date: "2/3/2022"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)

### for part 1: PCA
library(ggfortify)
library(palmerpenguins)
library(patchwork)

### for part 2: ggplot customization
library(readxl)
library(gghighlight)
library(lubridate)
library(plotly)
```

# Part 1: PCA (Principal Components Analysis)
A way to take a multivariate data set and get as much information as possible in a consice way. Gives you idea about the importance of variables. 

```{r}
penguin_clean <- penguins %>%
  drop_na()
# need to drop NA values or PCA analysis wont run

# create PCA object (contains a bunch of information within the model object that we can use for different purposes)
# we can only give this function NUMERIC data, so need to get rid of non-numeric variables from the ENTIRE  dataframe!!
penguin_pca <- penguin_clean %>%
  select(body_mass_g, ends_with('_mm')) %>% # ends_with selects all columns that end with... '_mm'
  scale() %>%
  prcomp()
           
# in console: View(penguin_pca) to see std deviations
penguin_pca$rotation # to see rotation values
penguin_pca$sdev #to see just std deviations
penguin_pca$x # where variables would fall in each principal component
```

## Biplot by hand

```{r}
loadings_df <- data.frame(penguin_pca$rotation) %>%
  mutate(axis = row.names(.)) # . = take the info from above and put it in the parentheses
# need row names, create new column

new_pts_df <- data.frame(penguin_pca$x)

ggplot() +
  geom_point(data = new_pts_df, aes(x = PC1, y = PC2), color = 'blue') +
  geom_segment(data = loadings_df, x = 0, y = 0,
               aes(xend = PC1, yend = PC2, group = axis),
               arrow = arrow(length = unit(.25, 'cm'))) +
  geom_text(data = loadings_df, aes(x = PC1, y = PC2, label = axis),
            vjust = 0, nudge_y = .02) +  #vjust = vertical justification of labels 0 = at the bottom of the arrow, nudge_y = nudge upward just a little
  theme_minimal()
```
### Biplot Automated
```{r}
autoplot(penguin_pca, 
         data = penguin_clean,
         loadings = TRUE, # will show arrows
         colour = "species", # have to spell colour because package is old
         loadings.label = TRUE,
         loadings.colour = 'black',
         loadings.label.colour = 'black',
         loadings.label.vjust = -.5) +
  scale_color_manual(values = c('blue', 'purple', 'orange')) + 
  theme_minimal()
```

## Screeplot by hand

```{r}
# make vector of std deviations from PCA
sd_vec <- penguin_pca$sdev

# need variance for screeplot so convert sd -> variance
var_vec <- sd_vec^2

pc_names <- colnames(penguin_pca$rotation)

# make percent of variance explained df
pct_expl_df <- data.frame(v = var_vec,
                          pct_v = var_vec/sum(var_vec),
                          pc = fct_inorder(pc_names)) %>%
  mutate(pct_lbl = paste0(round(pct_v * 100, 1), '%'))

ggplot(pct_expl_df, aes( x = pc, y = v)) +
  geom_col() + 
  geom_text(aes(label = pct_lbl), vjust = 0, nudge_y = .005) %>%
  labs(x= "Principal component", y = "Variance explained")
# vertical bars say HOW MUCH VARIANCE, Labels = what percent of the total variance is captured
  
```

## Screeplot Automated
```{r}
# use screeplot() to take PCA object and quickly graph into screeplot (quick version of above)
# can specifiy different types
screeplot(penguin_pca, type = "lines")
screeplot(penguin_pca, type = "barplot")
```


# Part 2: ggplot customization, reading in different file types

```{r}
# read in and clean data
fish_noaa <-read_excel(here("foss_landings.xlsx")) %>% 
  janitor::clean_names() %>% 
  mutate(across(where(is.character), tolower))  %>%# mutate across multiple columns where is.character is TRUE and make it lowercase (converts values to all lowercase if they are character value)
  mutate(nmfs_name = str_sub(nmfs_name, end = -4)) %>% # deletes ** from end of nmfs_name values
  filter(confidentiality == "public")
```

```{r}
fish_plot <- ggplot(data = fish_noaa, aes(x = year, y = pounds, group = nmfs_name)) +
  geom_line(aes(color = nmfs_name)) + 
  theme_minimal()
 
# interactive plot! 
ggplotly(fish_plot)


# gghighlight
ggplot(data = fish_noaa, aes(x = year, y = pounds, group = nmfs_name)) +
  geom_line() +
  gghighlight(nmfs_name == 'rockfishes') + # highlight based on rockfishes
  theme_minimal()

ggplot(data = fish_noaa, aes(x = year, y = pounds, group = nmfs_name)) +
  geom_line(aes(color = nmfs_name)) +
  gghighlight(max(pounds) > 1e8) + # this will highlight based on max catch
  theme_minimal()
```

## Read in data straight from the web!?! and practice lubridate
lubridate helps you more easily manage dates in different formats
```{r}
monroe_wt <- read_csv("https://raw.githubusercontent.com/oharac/esm244_w22_lab5/master/data/MWTP_Elec_Daily.csv") %>% janitor::clean_names()

# date format is not good, need to reformat:
monroe_ts <- monroe_wt %>%
  mutate(date = lubridate::mdy(date))  %>% # check using `class(monroe_ts$date)`
  mutate(record_month = lubridate::month(date), # new column with month number pulled from the date column
         month_name = month.abb[record_month]) %>% # month.abb = abbreviations of months, use [] to locate index
  mutate(month_name = fct_reorder(month_name, record_month)) # reorders month names in order of month number

ggplot(data = monroe_ts, aes(x = month_name, y = total_kwh)) +
  geom_jitter(aes(color = month_name),
              show.legend = FALSE,
              alpha = 0.5,
              size = 0.3,
              width = 0.2) +
  theme_classic()

```





